\chapter*{Acknowledgements}

\begin{onehalfspace}

This is the very last thing I'm writing before sending this to print, because I'm awful at writing these sort of things and I kept putting it off. Not that it really matters, of the very few people who will ever read this thesis I'm guessing most will skip over this bit. I always do.

\medskip

Anyway, I'll start by thanking my PhD supervisor Vik Dhillon, for being a fantastic supervisor and mentor over the past four years, as well as, when needed, a 5-star tour guide of La Palma. Thanks also to my second supervisor Ed Daw, whose Physical Computing class I enjoyed both teaching and learning from, as well as Stu, Dave, Steven, Liam, James, Lydia, Pablo and others in the group in Sheffield. I'd also like to thank members of the GOTO collaboration: Danny, Krzysztof, Paul, Joe, James and Ryan at Warwick; Duncan, Kendall, Evert, Travis and Alex at Monash; Gav, Mark and everyone else from the ever-increasing list of member institutions (especially the poor students who had to babysit GOTO for so long). Also my masters project supervisors at Durham, Richard and Tim, who first put me in touch with Vik --- without them I literally would not be where I am today.

\medskip

I'd like to thank all the various friends I've made over the years from Maidenhead, Borlase, Durham and Sheffield; including, but not at all limited to, Alex, Anna, Becky, Ben, Emma, Gemma, Harriet, Harry, Héloïse, James, Katie, Katherine, Laura, Mac, Oliver, Richard, Robin, Simon, Sophie and Varun (I wanted to include as many people as I could, apologies to anyone I forgot). A particular massive thanks to the other PhD students in my year at Sheffield: Becky, Héloïse and Katie. You were the best office-mates and compatriots I could have asked for, and while you all finished before me and went off to better things, it did mean I my own office in the final few weeks --- thanks!

\medskip

Of course, I must also thank my Mum and Dad, and my brother Phil, for their unwavering support and encouragement, as well as my Grandad Norman, my cousin Nye, and the rest of my family.

\newpage

\medskip

Finally, I would like to acknowledge the people who did the most to encourage my love of physics and astronomy that eventually culminated in this thesis: my secondary-school physics teacher Malcolm Brownsell, the members of the Maidenhead Astronomical Society, and my grandfather Eric Snelling, aka Bubba. May you all have clear skies.

\smallskip

\begin{flushright}
--- Martin Dyer, Sheffield, 25th September 2019
\end{flushright}

\medskip

PS Although the above is dated the 25th, and I did try to print out and submit on that day, I failed to do so --- the office didn't have any binding combs large enough for this epic tome. So I'm going to get it done today, the 26th, at the Rymans in town, and I'll leave this postscript as a permanent record. Let's hope today is the day!

\end{onehalfspace}
